---
title: "Mixed Layer Depth"
author: "David Kaiser"
date: "2021/08/29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, fig.height=5, fig.width=4, fig.align = "center", comment = NULL)
source("mixed_layer_depth_DK.R")
```

## Description

A function to calculate the depth of the surface mixed layer of a water column, defined as the depth at which a reference parameter changes beyond a given threshold relative to the first value. 

## Arguments

*depth* -- vector for depth (in distance or pressure), can be negative 

*variable* -- temperature or density

*threshold* -- the max difference of *variable*, relative to the surface, that marks the end of the MLD

*print.info* = FALSE -- should into be printed in the console as a side effect

*plot* = FALSE -- should the result be plotted as a side effect

*depth.max* = NULL -- maximum depth to be plotted, uses absolute value of max depth if no value is supplied


## Result

A numeric value of the Mixed Layer Depth. If the threshold is not exceeded the result will be NA. This is preferable to using the max depth as the data might not include the entire water column and might thus not reach the thermo/pycnocline. 

Former side effects of a plot showing the data and the resulting mixed layer depth (if any) and the threshold values, and a printed explanation of the result are now optional and off by default.

## Example

example data
```{r example data}
mixed <- data.frame(depth = seq(from=0, to=99, by=1), 
                    temperature = rep(c(7,7.1), times = 50))
stratified <- data.frame(depth = seq(from=0, to=99, by=1), 
                         temperature = c(rep(20, times=25), seq(from=20, to=7, length.out=25), seq(from=7, to=5, length.out=50)))

```

For a **stratified** water column
```{r stratified_example}
MLD_DK(depth = stratified$depth, variable = stratified$temperature, threshold = 0.5) 
```

For a **mixed** water column
```{r mixed_example}
MLD_DK(depth = mixed$depth, variable = mixed$temperature, threshold = 0.5) 
```

With **side effects**
```{r side_effects}
MLD_DK(depth = stratified$depth, variable = stratified$temperature, threshold = 0.5, 
              print.info = TRUE, plot = TRUE, depth.max = 50)
```

With **negative depth**
```{r neg_depth}
MLD_DK(depth = -stratified$depth, variable = stratified$temperature, threshold = 0.5, 
              print.info = TRUE, plot = TRUE)
```
